---
title: "Astellas metastatic PDAC"
subtitle: "Feasibility report"
author: "Alex Paynter"
date: "`r format(Sys.Date(), '%d %b %Y')`"
editor_options:
  quarto:
    chunk_output_type: console
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 4
    theme: sandstone 
execute:
  echo: false
  include: false
  warning: false
  message: false
  fig.width: 7
  fig.height: 5
---

```{r}
library(fs)
library(purrr)
library(here)
purrr::walk(.x = fs::dir_ls(here("R")), .f = source)
```


```{r}
dt_wrapper <- function(x, cap = NULL, font.size = '10pt', page_length = 10) {
  DT::datatable(
    x,
    style = "bootstrap4",
    fillContainer = F,
    rownames = F,
    caption = cap,
    options=list(
      initComplete = htmlwidgets::JS(
        "function(settings, json) {",
        paste0("$(this.api().table().container()).css({'font-size': '", font.size, "'});"),
        "}"),
      pageLength = page_length
    ) 
  )
}

theme_gtsummary_compact(font_size = 12)
theme_gtsummary_language("en", big.mark = "") 
```


## Introduction

This report addresses the feasibility and resultant cohort of an Astellas study of second line therapies in metastatic pancreatic ductal adenocarcinoma (PDAC).  We will leverage the existing curation in the pancreatic BPC (v1.2-consoritum) cohort to study the questions outlined.

## Attrition table

We'll show the table first, and detailed explanations follow below.  Each step in the table is cumulative, meaning that the second row states the number who meet the conditions in the first and second rows.

```{r}
flow_track <- readr::read_rds(
  here('data', 'flow_track.rds')
)

flow_track_ft <- flow_track %>%
  mutate(n = map_int(dat, nrow)) %>%
  select(message, n) %>%
  flextable() %>%
  set_header_labels(message = "Cohort description", n = "n") %>%
  theme_booktabs() %>%
  align(j = "n", align = "right") %>%
  fontsize(size = 10) %>%
  autofit() 
```

```{r}
#| include: true
flow_track_ft
```

Details:

- **BPC PANC v1.2** The biopharma collaborative pancreatic cancer dataset, version 1.2-consortium.
- **PDAC tumor** - At least one sample for each person had the oncotree code for Pancreatic Adenocarcinoma (PAAD) or Adenosquamous Carcinoma of the Pancreas (PAASC), leaving out anyone with only Acinar Cell Carcinoma of the Pancreas (PAAC) and Undifferentiated Carcinoma of the Pancreas (UCP) samples.
- **US sites only** Eliminated UHN (Canada) from the cohort.
- **Only one cancer** The only cancer these people had is the pancreatic cancer that qualifies them for the cohort.  Only about 30 people had multiple cancers with pancreatic first, so it's almost certainly not worth it to open the can of worms that comes with trying to figure out which therapy counts for which cancer.
- **Met dx. (anytime)** Limits to those who had a metastatic diagnosis at some point.  We will check the timing of the metastatic diagnosis with the index therapy later.
- **KRAS G12D+** Had the mutation of interest in at least one NGS test (see note below on timing).
- **1L gem/5FU-based** - The first line of therapy includes either gemcitabine or 5-FU (Fluorouracil).  Using the wording in inclusion criterion #3 from the protocol, lines of therapy start with this exposure even if the exposure is before metastasis.
- **Metastatic at 2L/3L** - the person is metastatic at the index therapy, which is either one or two therapies after the gem/5fu exposure (2L or 3L).
- **No investigational** - the person had no investigational agents included in any therapies from the gem/5-FU to the index therapy.
- **Documented prog after 1L** - complicated, see *Progression* section below.

```{r}
cpt_index_timing <- readr::read_rds(here('data', 'drug', 'cpt_index_timing.rds'))
n_rep_before_index <- cpt_index_timing %>%
  filter(pos_rep_before_index) %>%
  nrow
n_ord_before_index <- cpt_index_timing %>%
  filter(pos_ord_before_index) %>%
  nrow
```


*Note on NGS timing:* We had some questions about the timing of KRAS G12D+ results and index therapy start times.  Two numbers on this subject:

- The number of people who have their KRAS G12D+ test result before they begin their index therapy (and meet all other criteria in this list) is `r n_rep_before_index`.  
- The number who had a test ordered which would eventually come back positive is `r n_ord_before_index`.

```{r}
demo_dat <- readr::read_rds(
  here('data', 'baseline_char.rds')
)
```


## Demographics

The following table shows the demographics for the group meeting all inclusion criteria:

```{r}
#| include: true
gtsummary::tbl_summary(
  select(demo_dat, -record_id),
  digits = list(
    `Year of birth` ~ 0
  )
)
```



## Progression

Progression is defined as one of the following two events happening between 6 weeks (42 days) and 1 year (365 days) after the initiation of the gem/5-FU therapy:

- A medical oncology note documented a worsening, or an imaging report radiologist summary noted a worsening. 
- There was a change in therapies, and there were no evaluations by imaging or medical oncology documented up to the date of the switch.  This second condition contributed a tiny number of cases (approximately 3) and could be removed for simplicity if desired.

## Lines

We begin counting regimens as lines of therapy when either of the following happens:

1. Exposure to gemcitabine or 5-FU (even if they precede the metastatic diagnosis).
2. Any therapy given in the metastatic setting.

Additionally, we do not count consecutive regimens using the same drugs as a new line of therapy.  For example, if a patient's first three captured regimens are FOLFIRINOX, FOLFIRINOX, Gemcitabine + Nabpaclitaxel, then we would say their first line is FOLFIRINOX and the second line is Gemcitabine + Nabpaclitaxel.  That is, the second second administration of FOLFIRINOX is not a new line but the continuation of an old line.

If desired, we can expand this concept to say that similar regimens do not count as new lines.  For example, if someone goes from FOLFIRINOX to FOLFOX you could argue this dropping one for toxicity rather than a new line. We have not done this today because it requires detailed decisions.

```{r}
lot <- readr::read_rds(here('data', 'drug', 'lot.rds'))
cohort <- readr::read_rds(here('data', 'cohort_prog_verified.rds'))
```

The following table shows all the first line (gem/5-FU) therapies:

```{r}
#| include: true

lot %>%
  filter(record_id %in% cohort$record_id) %>%
  filter(line_of_therapy %in% 1) %>%
  count(regimen_drugs, sort = TRUE) %>%
  dt_wrapper(
    .,
    cap = "Regimens for the Gem/5-FU exposure (1L)",
    page_length = 10
  )
```

This table shows the therapies used in the index line (either 2L or 3L):

```{r}
#| include: true

cohort %>%
  select(record_id, line_of_therapy = index_line) %>%
  left_join(., select(lot, record_id, line_of_therapy, regimen_drugs),
            by = c('record_id', 'line_of_therapy')) %>% 
  count(regimen_drugs, sort = TRUE) %>%
  dt_wrapper(
    .,
    cap = "Regimens for index therapy (2L/3L)",
    page_length = 10
  )
```

**Note:** Lots of these are Gem and 5-FU, too, which may or may not be what the sponsor was expecting.


```{r}
met_sum <- readr::read_rds(
  here('data', 'met_site_sum.rds')
)
```

## Sites of metastasis

The following table shows the number (%) of patients with a distant metastasis at select body regions when they begin the index therapy.  As discussed, more detailed information (ICD-O-3 codes) on metastatic sites can be mined from the imaging data if needed.

```{r}
#| include: true
gtsummary::tbl_summary(
  select(met_sum, -c(record_id))
) %>%
  modify_header(label = "**Met at index therapy**")

```


## Important changes

- Jul 2025: Added a table of characteristics at index therapy.
- Jul 2025: The first draft of our progression algorithm allowed for progression defined by a drug switch, only in participants that had no medical oncologist or imaging assessments after starting first line therapy.  This has been removed (**Impact: ~-3 cases**).
- Jul 2025: Previously we only included progression events which occurred >=6 weeks and <=6 months after 1L therapy.  After learning this does not match the Astellas protocol we removed this (**Impact: ~+30 cases, OS longer**).
- Jul 2025: Originally progression was determined by either medical oncologist or imaging note.  We changed to just imaging to align with the target trial progression definition and the definition of response, which both use imaging only (**Impact:~-18 cases**).


## Miscellaneous notes

- Using oncotree codes, we counted PAAD and PAASC as PDAC while PAAC and UCP are not.  Here are the four types present in the data:
    - Pancreatic Adenocarcinoma (PAAD)
    - Adenosquamous Carcinoma of the Pancreas (PAASC)
    - Acinar Cell Carcinoma of the Pancreas (PAAC)
    - Undifferentiated Carcinoma of the Pancreas (UCP)
- The interval for progression could be modified if needed, ours is just a starting point for discussion.