---
title: "Astellas metastatic PDAC"
subtitle: "Data checking"
author: "Alex Paynter"
date: "`r format(Sys.Date(), '%d %b %Y')`"
editor_options:
  quarto:
    chunk_output_type: console
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 4
    theme: sandstone 
execute:
  echo: false
  include: false
  warning: false
  message: false
  fig.width: 7
  fig.height: 5
---
  

```{r}
library(fs)
library(purrr)
library(here)
purrr::walk(.x = fs::dir_ls(here("R")), .f = source)
```

## Introduction

This report lists out cases to be visually confirmed in cBioPortal.  The idea is that each link is a query on one person, and with a little bit of additional information (which drug is index) you can check that they meet the key entry and response criteria.

This report is intended for internal (Alex) use, but if you want to spend some time checking things over I'm fully supportive - reach out if you have problems.

```{r}
cohort <- readr::read_rds(
  here('data', 'cohort_prog_verified.rds')
)
resp <- readr::read_rds(
  here('data', 'response', 'responses.rds')
)

chk_dat <- cohort %>%
  select(
    record_id,
    index_drugs = regimen_drugs
  ) %>%
  left_join(
    .,
    select(resp, record_id, n_scans_eval, resp_cat, resp_cat_conf, bor),
    by = 'record_id'
  )

chk_dat %<>%
  mutate(
    chk_cat = case_when(
      resp_cat_conf ~ "con_res",
      resp_cat ~ "uncon_res",
      n_scans_eval == 0 ~ "ne_scans_only",
      !resp_cat ~ "non_res",
      T ~ "no_scans" # could also look for n_scans_eval being NA.
    ) 
  )

chk_dat %<>%
  mutate(
    cbio_q = output_cbio_lines(record_id, return_vec = T)
  )
```

```{r}
 # This needs some fixing.
line_outter <- function(dat) {
  for (i in 1:nrow(dat))
    cat(paste0(dat[[i, "record_id"]], "\n\n"),
        paste0("    - ", dat[[i, "index_drugs"]], "\n"),
        paste0("    - <", dat[[i, 'cbio_q']], ">\n\n"))
}
```


## Section 1: Response strata 

Categories:

- Confirmed responders
- Unconfirmed responders
- Non-responders
- NE scans only
- No scans

What I want to check:

- Everyone has a progression before their index line.
- Everyone's response category looks correct.

For each person below we write the subject ID, index drug line (because you can't see that in cBio), and a link to a cbio query for that person's clinical view.

### Confirmed responders

```{r}
#| output: true
#| include: true

set.seed(23048)

chk_dat %>%
  filter(chk_cat %in% "con_res") %>%
  dplyr::sample_frac(1) %>%
  line_outter
```


### Unconfirmed responders

```{r}
#| output: true
#| include: true

chk_dat %>%
  filter(chk_cat %in% "uncon_res") %>%
  dplyr::sample_frac(1) %>%
  line_outter
```


### Non-responders

```{r}
#| output: true
#| include: true

chk_dat %>%
  filter(chk_cat %in% "non_res") %>%
  dplyr::sample_frac(1) %>%
  line_outter
```


### No scans

```{r}
#| output: true
#| include: true

chk_dat %>%
  filter(chk_cat %in% "no_scans") %>%
  dplyr::sample_frac(1) %>%
  line_outter
```

### NE scans only

```{r}
#| output: true
#| include: true

chk_dat %>%
  filter(chk_cat %in% "ne_scans_only") %>%
  dplyr::sample_frac(1) %>%
  line_outter
```


## Other stuff

### non-Progressers?

