---
title: "Astellas metastatic PDAC"
subtitle: "Real world response report"
author: "Alex Paynter"
date: "`r format(Sys.Date(), '%d %b %Y')`"
editor_options:
  quarto:
    chunk_output_type: console
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 4
    theme: sandstone 
execute:
  echo: false
  include: false
  warning: false
  message: false
  fig.width: 7
  fig.height: 5
---


```{r}
library(fs)
library(purrr)
library(here)
purrr::walk(.x = fs::dir_ls(here("R")), .f = source)
```

## Introduction

This report details the real-world response data available in GENIE BPC.  GENIE is a real-world study, so the statistics in this report should not be expected to follow RECIST.  Specifically, in radiology reports where cancer is detected we have the following options describing the change in status (the **[bold]** annotations represent a shorthand for this study):

> Radiologist Assessment of Change in Cancer Status [image_overall]
>
> • Improving/Responding **[PR]**
> 
> • Stable/No change **[SD]**
>
> • Mixed **[MR]**
>
> • Progressing/Worsening/Enlarging **[PD]**
> 
> • Not stated/Indeterminate **[NE]**


Everything response determination in this report is based on this information.

RECIST demands that the sum of tumor diameters decrease by at least 30% and 5mm in order to note a response.  Practices in the real world vary, but generally we expect that radiologist impressions will note improvements smaller than this as "Improving/Responding".  We have left the outcome titles as is (that is, PR instead of rwPR and ORR instead of rwRR) to avoid clutter - everything here is real world. 

Followup for response starts the day after initiation of the index therapy.  Followup is terminated at death or end of followup (naturally) or a switch in therapy to the next line.  By request, we truncate follow-up at first progression, in order to match the existing trial arm criteria.  This does differ from the recommendations in the RECIST paper.

**Responders** are those with at least one radiology impression noting Improving/Responding (PR).  **Non-responders** are those with at least one radiographic asssessment but no indication of response.  **Best overall response (BOR)** shows the best response in the followup window, using the order in the levels above (that is, PD is only noted if MR is NOT noted).

Confirmed response requires the following sequence of events:

- *(optional)* 0+ SD or NE scans.
- **1 PR scan.**
- *(optional)* 0+ NE scans.
- **1 SD or PR scan.**

The last line follows because we expect real-world impressions to reference the *previous* scan, so SD after PR most likely represents an improvement from baseline.  This is discussed on page 13 of the supplement in the "rwSD" definition in <https://doi.org/10.1200/CCI.24.00091>.

The denominator for response rate will be those with at least one radiographic scan after index therapy, even for confirmed response calculations.  The justification is again matching the trial arm protocol, but it's easy to add the people no scans into the denominator by hand if needed.

```{r}
cohort <- readr::read_rds(
  here('data', 'cohort_prog_verified.rds')
)
resp <- readr::read_rds(
  here('data', 'response', 'responses.rds')
)
```



## Results

There were `r nrow(cohort)` people in our cohort which met eligibility criteria.  Of those, `r nrow(resp)` had at least one radiographic scan after initiation of index therapy, forming the denominator for ORR analyses below.  



```{r}
incl_img <- readr::read_rds(
  here('data', 'response', 'incl_img.rds')
)
incl_img %<>%
  mutate(
    index_scan_days = dob_event_days - dob_index_start_int
  ) 

first_scans <- incl_img %>%
  group_by(record_id) %>%
  summarize(
    first_scan = min(index_scan_days, na.rm = T)
  )

first_scans_eval <- incl_img %>%
  group_by(record_id) %>%
  filter(!is.na(status_change)) %>%
  summarize(
    first_eval_scan = min(index_scan_days, na.rm = T)
  )

first_scans <- full_join(
  first_scans,
  first_scans_eval,
  by = 'record_id'
)

  
  
```



### Placeholder:  Stats on time to scans

We will do this with KM methods later.  For now we offer the empirical stats describing only those who had scans and contribute to the numbers below:

```{r}
#| include: true
first_scans %>%
  select(
    `Time to first scan (days)` = first_scan,
    `Time to first scan with status (days)` = first_eval_scan) %>%
  gtsummary::tbl_summary(.)
```

Unknown above is the people who have only not-evaluable scans after index therapy.






```{r}

resp_gt <- resp %>%
  mutate(bor = factor(bor,
                      levels = c("PR", "SD", "MR", "PD", "NE"))) %>%
  select(
    n_scans,
    n_scans_eval,
    `Best overall response` = bor
  ) %>%
  gtsummary::tbl_summary(.) 
```




### Best overall response + scan stats

The following table shows the number of scans, number of evaluable scans and the best overall response.  The priority order on BOR is the order of the table, so anyone with a "PR" response in the followup window would not be classed into a lower category (for example).

```{r}
#| include: true

resp_gt
```





```{r}
resp_ci <- binom::binom.confint(
  c(sum(resp$resp_cat), sum(resp$resp_cat_conf)),
  rep(nrow(resp), 2),
  methods = 'exact'
) %>%
  mutate(
    type = c('Any', 'Confirmed')
  )

resp_ci_ft <- resp_ci %>%
  mutate(
    `Non-responders` = n-x,
    `Rate (95% CI)` = glue('{formatC(mean*100, digits = 1, format = "f")}% ({formatC(lower*100, digits = 1, format = "f")}, {formatC(upper*100, digits = 1, format = "f")})')
    
  ) %>%
  rename(Responders = x) %>%
  select(
    ` ` = type, Responders, `Non-responders`, `Rate (95% CI)`
  ) %>%
  as_tibble(.) %>%
  flextable(.) %>%
  autofit(.)
    
```


### ORR rate

The following table shows the number of responders and non-responders, along with the rate and a 95% exact confidence interval.  This is repeated for any responder and only including those with a confirmed response (using the above definition).

```{r}
#| include: true
resp_ci_ft
```


### Placeholder: Discussion on RECIST alignment for those 0-30%




## Appendix

```{r}
resp_code_binning_dt <- resp %>%
  count(
    resp_codes_trunc,
    resp_cat,
    resp_cat_conf,
    bor,
    sort = T,
    name = "n_cases"
  ) %>%
  mutate(
    resp_codes_trunc = unroll_letter_codes(resp_codes_trunc)
  ) %>%
  select(
    n_cases,
    Scans = resp_codes_trunc,
    Responder = resp_cat,
    `Conf. Resp` = resp_cat_conf,
    BOR = bor
  )

resp_code_binning_dt %<>%
  dt_wrapper(.)
    
```


### Response categorization

The following table shows every unique sequence of scans experienced by patients.  For each unique sequence we show responder (true/false), confirmed responder (true/false) and BOR.  The goal of this table is facilitating a detail-oriented discussion on some of the edge cases which come up, limited to those which actually occurred.

```{r}
#| include: true

resp_code_binning_dt

```


```{r}
resp_code_trunc_dt <- resp %>%
  count(
    resp_codes, 
    resp_codes_trunc,
    sort = T,
    name = "n_cases"
  ) %>%
  mutate(
    resp_codes = unroll_letter_codes(resp_codes),
    resp_codes_trunc = unroll_letter_codes(resp_codes_trunc),
    diff = resp_codes != resp_codes_trunc
  ) %>%
  select(
    n_cases,
    Raw = resp_codes,
    Truncated = resp_codes_trunc,
    diff
  ) %>%
  arrange(desc(diff), desc(n_cases))

resp_code_trunc_dt %<>%
  dt_wrapper(.)
```

### Truncation of followup

As stated in the introduction, followup was truncated at the first sign of progression.  The following table shows the Truncated scan sequence and the Raw scan sequence, to show the impact of this.  The overall impact is one directional:  The number of responders will decrease when truncating followup.

By default the table shows the cases where the truncation makes a difference in the included scans (diff = true) at the top, though this is not always consequential (actually, usually not).

```{r}
#| include: true
resp_code_trunc_dt
```




## Questions

- Mixed response - should it be considered equivalent to SD?  Affects both confirmed response definition and BOR definition.  Currently mixed response effectively acts like a progression, not counting as either a confirmation of response or as "padding" while waiting for something more definitive.
- Should those people with only non-evaluable scans after index therapy be included in the denominator for ORR?  Currently they are.  It's only 3 people, but I think this comes down to the trial (which may not have NE scans).
- In the case where a person switches therapy, we count any scans done on that day.  For example, if someone started a therapy at day 0, and switched therapy on day 100, we still consider a scan on day 100 to be indicative of response/not to the therapy started on day 0.  In fact, it seems most plausible to me that the switch was prompted by a lack of response recorded on that day.  This does actually affect the inclusion of 4 scans, which may or may not alter the results.