---
title: "Astellas metastatic PDAC"
subtitle: "Survival report"
author: "Alex Paynter"
date: "`r format(Sys.Date(), '%d %b %Y')`"
editor_options:
  quarto:
    chunk_output_type: console
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 4
    theme: sandstone 
execute:
  echo: false
  include: false
  warning: false
  message: false
  fig.width: 7
  fig.height: 5
---

```{r}
library(fs)
library(purrr)
library(here)
purrr::walk(.x = fs::dir_ls(here("R")), .f = source)
```

## Introduction

This report examines overall survival in the cohort detailed in the `cohort_build` report.  These people meet the criteria of the Astellas study to the extent that we're able to check/match them.

This cohort is now limited to participants who have an NGS result back indicating KRAS G12D+ before the start of the index therapy.  By definition, this eliminates anyone with delayed entry, so earlier discussions of left truncation are not longer relevant.  The other expected effects of this change are:

  - Smaller cohort
  - People who have not received an NGS test by 2L metatstatic disease are more likely to be experience a lower standard of care.  We cannot confirm this empirically in GENIE.
  - As a result of the above, the OS estimates here are longer.

There is a section which provides risk set adjusted analyses (on the larger cohort which may or may not have NGS resulted at the time of index therapy), showing the effect of this change.

```{r}
main_model_bundle <- readr::read_rds(
  here('data', 'survival', 'main_model_bundle.rds')
)
geno_compare_bundle <- readr::read_rds(
  here('data', 'survival', 'geno_comp_bundle.rds')
)
```

## Overall survival

The primary analysis for this stage of the study is a one-arm KM estimate of overall survival (95% CI).  There are `r nrow(main_model_bundle$dat_surv)` cases which contribute risk time to this analysis.  The following table and figure show these numbers, along with a table of survival percentages at some select times:

```{r}
#| include: true
#| fig-height: 4

main_model_bundle$gg_os
```

```{r}
#| include: true
main_model_bundle$gt_med_surv
```


```{r}
#| include: true
main_model_bundle$gt_surv_times
```



## Genomic OS comparisons

We received a request to make two comparisons if time permitted:

- Those with KRAS G12D mutations, compared to those who only have other G12 mutation ("G12[*]").
- Those with KRAS G12D mutations, compared to those who have no G12 mutations of any kind.  We will denote this as "no G12D", as they could have other KRAS mutations.

These are shown below, with little difference in survival based on genomics.  These cohorts are built in the same way as the main analysis, just skipping the KRAS G12D+ requirement.  The tables show 95% confidence intervals for median survival.

```{r}
#| include: true

geno_compare_bundle$gg_os_geno_1
geno_compare_bundle$gt_median_surv_1
```

```{r}
#| include: true

geno_compare_bundle$gg_os_geno_2
geno_compare_bundle$gt_median_surv_2
```


## Progression-free survival

In GENIE PFS can be calculated by requiring progression by imaging, med onc notes, either or both.  Owing to our previous conversations that the Astellas trial is an imaging-based endeavor, our calculation here is based on imaging only.  In general we would recommend the use of PFS I-and-M (requiring both imaging and medical oncologist assessement for progression), due to the superior correlation with overall survival.  This is discussed for NSCLC in <https://doi.org/10.1001/jamanetworkopen.2021.17547>.


```{r}
#| include: true
#| fig-height: 4

main_model_bundle$gg_pfs
```

```{r}
#| include: true
main_model_bundle$gt_med_pfs
```


```{r}
#| include: true
main_model_bundle$gt_pfs_times
```

